---
title: Methods for Risk Evaluation part 2
author: Jingchen (Monika) Hu 
institute: Vassar College
date: Statistical Data Privacy
output:
  beamer_presentation:
    includes:
      in_header: ../LectureStyle.tex
slide_level: 2
fontsize: 11pt
bibliography: ../book.bib

---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})
```

```{r message = FALSE, echo = FALSE, warning = FALSE}
require(smoothmest)
require(dplyr)
require(readr)
require(ggplot2)
require(coda)
require(gridExtra)
require(brms)
require(bayesplot)
```

## Outline

\tableofcontents[hideallsubsections]


# Introduction

## Recap

- Lecture 8:
    - Identification disclosure
    - Matching-based method
    

## Plan for this lecture

- Two general types of disclosure: identification and attribute (Lecture 1)

1. Identification disclosure: The intruder correctly identifies records of interest in the released synthetic data

2. Attribute disclosure: The intruder correctly infers the true confidential values of the synthetic records using information from the released synthetic data


- In this lecture, we focus on identification risk evaluation methods, with illustrations to the synthetic CE from Lectures 4 \& 5 and the synthetic ACS from Lecture 5


## Overview

- Identification disclosure: 
    - The intruder correctly identifies records of interest in the released synthetic data
    - Only exist in partially synthetic data

- We will introduce two general approaches
    - Matching-based approaches (last lecture)
    - Record linkage approaches

# Record linkage approaches

## Overview

- Record linkage methods are developed mainly for the purpose of linking records from multiple databases

- Based on variables, called **keys**, a link between two records can be established. Therefore, record linkage approaches can be used as metrics of identification risks (@Winkler2004PSD)

## Record linkage approaches for synthetic data

- For partially synthetic data, record linkage methods can be applied to linking records in the synthetic dataset to the records in the confidential dataset

- Among these linkages, we can evaluate identification risks in terms of true links (i.e., correct links) and false links (i.e., incorrect links)
    - high percentage of true links / low percentage of false links indicates high identification disclosure risk, and vice versa

## Record linkage approaches for synthetic data

- We now present the general procedure of performing probabilistic record linkage (@FellegiSunter1969JASA) to evaluate identification disclosure risks in partially synthetic data

- As with matching-based approaches
    - $\mathbf{Y} = (\mathbf{Y}^A, \mathbf{Y}^U)$ to represent the confidential data sample containing $n$ observations and $r$ variables
    - $\mathbf{Y}^A$ denotes the variables available to the intruder from external databases and $\mathbf{Y}^U$ denotes the variables unavailable to the intruder
    - similarly, we have $\mathbf{Z} = (\mathbf{Z}^A, \mathbf{Z}^U)$ for a partially synthetic dataset of $\mathbf{Y}$
    - we can further split $\mathbf{Z}^A$ into $\mathbf{Z}^{A_s}$, the synthesized variables and $\mathbf{Z}^{A_{us}}$, the unsynthesized variables

## Procedure

1. Given $\mathbf{Y}^A$, the set of variables available to the intruder, we generate pairs between $\mathbf{Y}$ and $\mathbf{Z}$ based on $\mathbf{Y}^A$ and $\mathbf{Z}^A$
    - that is, a pair of record $i$ from $\mathbf{Y}$ and record $j$ from $\mathbf{Z}$ is generated only when $\mathbf{y}_i^A = \mathbf{z}_j^A$ (for the entire vector)
    - note that since $\mathbf{Z}^A = (\mathbf{Z}^{A_s}, \mathbf{Z}^{A_{us}})$, some of these pairs would be incorrect, thanks to the changes brought by the data synthesis process
    - call this collection of pairs as $P$.

2. For each pair of records in $P$, we, the data disseminators, compare the values of the unavailable variables
    - for example, if $\mathbf{y}_i$ and $\mathbf{z}_j$ is paired up in step 1, then this step compares the values of $\mathbf{y}_i^U$ and $\mathbf{z}_j^U$
    - we can create a set of similarity score over all the unavailable variables, which will be used for scoring all pairs next.

## Procedure

3. With calculated similarity score for each pair, we then score all the pairs
    - think of this as ranking all pairs
    - the higher the ranking one pair is, the more likely a link will be established
    - this is the core of probabilistic record linkage as proposed by @Winkler2000Census, where we will use an expectation-maximization algorithm (**EM** algorithm)
    - in this approach, a weight value will be estimated for each pair, which will then be used next for determining links, also known as selecting pairs

## Procedure

4. Each pair now comes with a weight from step 3. We then select one-to-one linkages between records in $\mathbf{Y}$ and records in $\mathbf{Z}$
    - that is, each record in the synthetic dataset $\mathbf{Z}$ will be linked to at most one record in the confidential $\mathbf{Y}$, and vice versa

5. Among the selected pairs from step 4, we calculate the percentages of true links (i.e., the one-to-one links that are correct links) and of false links (i.e., the one-to-one links that are incorrect links)

## Example of the ACS sample

- Now we will illustrate this record linkage approach to evaluating identification disclosure risks in the synthetic ACS sample

- The probabilistic record linkage algorithm is implemented by the ```reclin``` package (@reclin)

```{r size = "footnotesize", warning = FALSE, message = FALSE}
## make sure to load the reclin package
library(reclin)
```

## Example of the ACS sample

We use the record linkage approach to evaluate the identification disclosure risks of a synthetic ACS sample in Lecture 5, where ```DIS, HICOV``` are synthesized and the other variables remain unsynthesized. The synthesis model is the DPMPM model with the *NPBayesImputeCat* R package. We assume assume that the intruder knows ```SEX, RACE, MAR``` of reach record. 

## Example of the ACS sample

- Load datasets

```{r size = "footnotesize", warning = FALSE, message = FALSE}
ACSdata <- data.frame(readr::read_csv(file = "ACSdata.csv"))
n <- dim(ACSdata)[1]
ACSdata_syn <- data.frame(readr::read_csv(file = "ACSdata_syn.csv"))
## make sure variables are in the same ordering
ACSdata_syn <- ACSdata_syn[, names(ACSdata)]
## add index for each record
ACSdata$id <- 1:n
ACSdata_syn$id <- 1:n
```

## Example of the ACS sample: generate pairs given available variables

- We first generate pairs given the available variables ```SEX, RACE, MAR```

- This is done by using the ```pair_blocking()``` function in the ```reclin``` package, where the inputs include ```ACSdata_syn```, ```ACSdata```, and the set of available variables

- The collection of pairs is stored in ```ACS_pairs```

```{r size = "footnotesize", warning = FALSE, message = FALSE}
ACS_pairs <- reclin::pair_blocking(ACSdata_syn, ACSdata, 
                                   c("SEX", "RACE", "MAR"))
```

## Example of the ACS sample: compare pairs based on unavailable variables

- Next, for each generated pair in ```ACS_pairs```, we create a set of similarity score over all the unavailable variables: the synthesized ```DIS, HICOV``` and the unsythesized ```LANX, WAOB, MIG, SCH, HISP```

- For this step, we use the ```compare_pairs()``` function in the ```reclin``` package and we use ```jaro_winkler``` similarity score (@reclin)

```{r size = "footnotesize", warning = FALSE, message = FALSE}
ACS_pairs_keys <- reclin::compare_pairs(ACS_pairs, by = c("LANX", "WAOB", 
                                                          "DIS", "HICOV", 
                                                          "MIG", "SCH",  
                                                          "HISP"),
                                        default_comparator = 
                                          jaro_winkler(0.9))
```

## Example of the ACS sample: compare pairs based on unavailable variables

- For illustration purpose, we print out the first few rows of ```ACS_pairs_keys```, where ```x``` and ```y``` are the record indexes from the two datasets, and each variable column shows a similarity score for that pair on that variable. 

```{r size = "footnotesize", warning = FALSE, message = FALSE}
ACS_pairs_keys[1:3, ]
```


## Example of the ACS sample: score all pairs with EM and produce weights

- We then use the probabilistic record linkage approach with an EM algorithm to produce a weight for each scored pair

- The higher the weight is, the more likely the pair of records belong to the same record

- The EM algorithm is implemented using the ```problink_em()``` function and the weights are then calculated using the ```score_problink()``` function in the ```reclin``` package

```{r size = "footnotesize", warning = FALSE, message = FALSE}
m <- reclin::problink_em(ACS_pairs_keys)
ACS_pairs_keys_pRL <- reclin::score_problink(ACS_pairs_keys, 
                                             model = m, 
                                             var = "weight")
```

## Example of the ACS sample: score all pairs with EM and produce weights

- The printout of the first few rows of ```ACS_pairs_keys_pRL``` shows an additional column, ```weight```, which corresponds to the calculated weights for all pairs after the EM algorithm for the probabilistic record linkage procedure

```{r size = "footnotesize", warning = FALSE, message = FALSE}
ACS_pairs_keys_pRL[1:3, ]
```

## Example of the ACS sample: select one-to-one linkag

- Now with calculated weight for each pair, we perform a one-to-one linkage by comparing weights for all pairs while making sure that one record from the synthetic ```ACSdata_syn``` can be linked to at most one record from the confidential ```ACSdata``` and vice versa

- There are a few choices provided by the ```reclin``` package, and for our data size, we choose to use the greedy algorithm with the ```select_greedy()``` function 

```{r size = "footnotesize", warning = FALSE, message = FALSE}
ACS_pairs_keys_pRL <- reclin::select_greedy(ACS_pairs_keys_pRL, "weight", 
                                            var = "greedy", threshold = 0)
```

## Example of the ACS sample: select one-to-one linkag

- This process adds one more column ```greedy``` to ```ACS_pairs_keys_pRL``` which shows TRUE / FALSE: TRUE indicates a link and FALSE indicates no link

```{r size = "footnotesize", warning = FALSE, message = FALSE}
ACS_pairs_keys_pRL[1:3, ]
```

## Example of the ACS sample: calculate percentages of true links and false links

- Lastly, we need to evaluate among all the TRUE's in ```greedy```, how many of them are true links and how many are false links

- A true link refers to a correct linkage, i.e., record $i$ in ```ACSdata_syn``` is correctly linked to record $i$ in ```ACSdata```

- A false link refers to an incorrect linkage, i.e., record $i$ in ```ACSdata_syn``` is incorrectly linked to record $j$ in ```ACSdata``` where $i \neq j$

- To do so, we create a new column ```true``` by comparing the ID's of each pair

- We add the ID's from ```ACSdata_syn``` and those from ```ACSdata``` using the ```add_from_x()``` and ```add_from_y()``` functions respectively, and then compare if they are the same

```{r size = "footnotesize", warning = FALSE, message = FALSE}
ACS_pairs_keys_pRL <- add_from_x(ACS_pairs_keys_pRL, id_x = "id") 
ACS_pairs_keys_pRL <- add_from_y(ACS_pairs_keys_pRL, id_y = "id") 
ACS_pairs_keys_pRL$true <- ACS_pairs_keys_pRL$id_x == 
  ACS_pairs_keys_pRL$id_y
```

## Example of the ACS sample: calculate percentages of true links and false links

- Lastly, we tabulate the ```true``` and ```greedy``` columns, as below.

```{r size = "footnotesize", warning = FALSE, message = FALSE}
table(ACS_pairs_keys_pRL[c("true", "greedy")])
```

\textcolor{blue}{Discussion question}: What do the results show us?

## Example of the ACS sample: results of the confidential data

- See hidden R scripts to evaluate the results on the confidential data

- The true linkage percentage is $6458 / 10000 = 64.58\%$, and the false linkage percentage is therefore $3542 / 10000 = 35.42\%$

\textcolor{blue}{Discussion question}: How do the synthetic data provide privacy protection compared to the confidential data?

```{r size = "footnotesize", warning = FALSE, message = FALSE, echo = FALSE}
ACS_pairs_con <- reclin::pair_blocking(ACSdata, ACSdata, c("SEX", "RACE", "MAR"))

ACS_pairs_keys_con <- reclin::compare_pairs(ACS_pairs_con, 
                                            by = c("LANX", "WAOB", "DIS", "HICOV", "MIG", "SCH", "HISP"),
                                            default_comparator = jaro_winkler(0.9))

m_con <- reclin::problink_em(ACS_pairs_keys_con)
ACS_pairs_keys_pRL_con <- reclin::score_problink(ACS_pairs_keys_con, model = m_con, var = "weight")

ACS_pairs_keys_pRL_con <- reclin::select_greedy(ACS_pairs_keys_pRL_con, 
                                                "weight", var = "greedy", threshold = 0)

ACS_pairs_keys_pRL_con <- add_from_x(ACS_pairs_keys_pRL_con, id_x = "id") 
ACS_pairs_keys_pRL_con <- add_from_y(ACS_pairs_keys_pRL_con, id_y = "id") 
ACS_pairs_keys_pRL_con$true <- ACS_pairs_keys_pRL_con$id_x == ACS_pairs_keys_pRL_con$id_y

table(ACS_pairs_keys_pRL_con[c("true", "greedy")])
```

## Final comments

- Note that in our illustration, all the available variables, ```SEX, RACE, MAR```, are unsynthesized, so our first step of generating pairs would have no errors

- It is possible that the intruder's knowledge of available variables includes some synthesized variables, which means the first step would generate incorrect pairs

# Summary and References

## Summary

- Record linkage approaches for identification disclosure risk evaluations
    - the ```reclin``` R package
    - the true linkage percentage and the false linkage percentage
    
\pause

- No homework! But you should be working on disclosure risk evaluation for your project

- Lecture 10: Methods for risk evaluation part 3
    - @BaillargeonCharest2020PSD (CAP statistic)
    
    
## References {.allowframebreaks} 

